---
title: "Wimbledon 2023: Men's Singles and Women's Singles"
---
Wimbledon is considered the most prestigious tournament in tennis, and this year, it takes place from July 3rd to July 16th. For this project, I tried to predict who would win the Men's singles draw and the Women's singles draw. The data for this project was sourced from a sports betting website that included files for every official ATP or WTA match from 2000 onwards. Once files for 2021, 2022, and 2023 were obtained, each desired statistic was counted for each player by year and surface. Then, they were combined in an intuitive manner to create a probability of winning the tournament. With a fairly high confidence interval, the three top contenders for Men's singles are Alcaraz (~8%), Medvedev (~6.5%), and Djokovic (~6.5%), and the top contender for Women's singles is Swiatek (~9%).
```{r}
#| echo: false
#| message: false
#| label: setup
library(tidyverse)
library(readxl)
library(rstanarm)
library(rayshader)
library(rgl)
```

## Men's
```{r}
#| echo: false
#| warning: false

pregraph1 <- read_csv("pregraph.csv")

pregraph1$name <- factor(pregraph1$name, levels = c("alcaraz c.", "medvedev d.", "djokovic n.", "sinner j.", "rublev a.", "norrie c.", "tsitsipas s.", "fritz t.", "ruud c.", "zverev a.")) 

tennisplot <- pregraph1 |> 
  mutate(probability = as.numeric(probability)) |>
  mutate(segment = cut(probability,
                     breaks = c(-Inf, seq(-0.1, 0.2, 0.01), Inf),
                     labels = seq(-0.11, 0.2, 0.01))) |>
  summarize(total = n(), 
            .by = c(name, segment)) |> 
  mutate(name = as.factor(name)) |> 
  mutate(segment = as.factor(segment)) |> 
  ggplot() +
  geom_tile(aes(x = reorder(name, segment), y = segment, fill = total), 
            width = 0.9, height = 0.9) +
  scale_fill_gradient(low = "lightblue", high = "blue") +
  scale_color_continuous(limits = c(0, 1000)) +
  theme_classic() +
  scale_x_discrete(labels = c("C. Alcaraz", "D. Medvedev", "N. Djokovic", "J. Sinner", "A. Rublev", "C. Norrie", "S. Tsitsipas", "T. Fritz", "C. Ruud", "A. Zverev")) +
  scale_y_discrete(breaks = c("-0.09", "-0.06", "-0.03", "0", "0.03", "0.06", "0.09", "0.12", "0.15", "0.18")) +
  labs(x = "", y = "Probability",
       title = "Men's Singles Top Ten Front Runners (Expected Posterior Distribution)",
       subtitle = "Alcaraz, Medvedev, and Djokovic have the highest chances of winning.") +
  theme(legend.position = "none") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

tennisplot

#3D plot:
#plot_gg(tennisplot,
        #width = 5.5,
        #zoom = 0.65,
        #theta = 25,
        #phi = 50,
        #sunangle = 225,
        #soliddepth = -0.5,
        #raytrace = FALSE,
        #windowsize = c(2048,1536))
#rglwidget()
#options(rgl.useNULL = TRUE)
#rgl.close()

#render_snapshot()
```

This graph only takes into account matches from 2021-2023 which excludes some important data. For example, the majority of Djokovic's historic 20-year career that includes 23 Grand Slam titles - seven of which are Wimbledon titles. Additionally, because it only accounts for the past 3 years, it weighs Medvedev's recent-ish success more heavily than it should. In reality, Alcaraz and Djokovic have an about equal shot of winning Wimbledon this year, with Medvedev being lower than second. 

## Women's

```{r}
#| echo: false
#| warning: false
#| label: stats_women

```

```{r}
#| echo: false
#| warning: false

pregraphw1 <- read_csv("pregraphw.csv")

pregraphw1$name <- factor(pregraphw1$name, levels = c("swiatek i.", "sabalenka a.", "jabeur o.", 
                                                    "ostapenko j.", "rybakina e.", "kvitova p.", 
                                                    "garcia c.", "pegula j.", "bencic b.", 
                                                    "alexandrova e."))

tennisplotw <- pregraphw1 |> 
  mutate(probability = as.numeric(probability)) |>
  mutate(segment = cut(probability,
                     breaks = c(-Inf, seq(-0.1, 0.2, 0.01), Inf),
                     labels = seq(-0.11, 0.2, 0.01))) |>
  summarize(total = n(), 
            .by = c(name, segment)) |> 
  mutate(name = as.factor(name)) |> 
  mutate(segment = as.factor(segment)) |> 
  ggplot() +
  geom_tile(aes(x = name, y = segment, fill = total), width = 0.9, height = 0.9) +
  scale_fill_gradient(low = "thistle", high = "darkmagenta") +
  scale_color_continuous(limits = c(0, 1000)) +
  scale_x_discrete(labels = c("I. Swiatek", "A. Sabalenka", "O. Jabeur", "J. Ostapenko", "E. Rybakina", "P. Kvitova", "C. Garcia", "J. Pegula", "B. Bencic", "E. Alexandrova")) +
  scale_y_discrete(breaks = c("-0.12", "-0.09", "-0.06", "-0.03", "0", "0.03", "0.06", "0.09", "0.12", "0.15", "0.18", "0.21")) +
  theme_classic() +
  labs(x = "", y = "Probability",
       title = "Women's Singles Top Ten Front Runners (Expected Posterior Distribution)",
       subtitle = "Swiatek has the highest chance of winning.") +
  theme(legend.position = "none") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

tennisplotw
```

This graph excludes the full match history of almost every player in this draw, so it almost completely excludes players like Svitolina who didn't play on tour in the past few years due to injury/personal reasons. Other than that, I think it is fairly accurate.